Enzo: An Adaptive Mesh Refinement Code for Astrophysics

### 2013-05-30 numerical-conduction.tex
Reynolds edits.
Edited conduction section

\enzo\ implements the equations of isotropic heat conduction in a manner
similar to that of \citet{2007ApJ...664..135P}, where we model the
flux of isotropic heat transport:

\begin{equation}
\mathbb{F} = -\kappa_{sp} \nabla T
\end{equation}

Where $\kappa_{sp}$ is the Spitzer conduction coefficient, $4.6 \times 10^{-7}$~T$^{5/2}$ erg
s$^{-1}$~cm$^{-1}$~K$^{-1}$ \citep{1962pfig.book.....S}, and using a
value for the Coulomb logarithm, $\log \Lambda = 37.8$, that is
appropriate for the intracluster medium \citep{1988xrec.book.....S}.
It is quite possible that the local heat flux can become unphysical in
the high-temperature, low-density cluster regime when using this
formulation; thus, we take into account the saturation of the heat
flux at a maximum level of

\begin{equation}
j_{sat} \simeq 0.4 n_e k_b T \left( \frac{2 k_b T}{\pi m_e} \right)^{1/2}
\end{equation}

\citep{1977ApJ...211..135C}, and to ensure a smooth transition between
the Spitzer regime and the saturated regime we define an effective
conductivity using the formalism of \citet{1988xrec.book.....S}

\begin{equation}
\kappa_{eff} = \frac{\kappa_{sp}}{1 + 4.2 \lambda_e / \ell_T} \; ,
\end{equation}

where $\lambda_e$ is the electron mean free path and $\ell_T \equiv T
/ |\nabla T|$ is the characteristic length scale of the local
temperature gradient.  We also assume that the conductivity of the
plasma can be described in terms of an effective conductivity, which
can be expressed as a fraction f$_{Sp}$ of the Spitzer conductivity
(where f$_{Sp} \leq 1.0$ are considered physical values).  This takes
into account physical processes below the resolution limit of the
simulation (such as tangled magnetic fields) that can suppress heat
transport.

Thermal conduction in a plasma can be strongly affected by the
presence of magnetic field lines, which may strongly suppress heat
flow perpendicular to the magnetic field.  In that case, we allow for
heat transport only parallel to the magnetic field lines in
magnetohydrodynamic simulations.  Mathematically, this is treated as:

\begin{equation}
\mathbb{F} = -\kappa_{sp} \hat{b} (\hat{b} \cdot \nabla T)
\end{equation}

Where $\hat{b}$ is the unit vector denoting the direction of the
magnetic field.  As with the isotropic thermal conduction, we allow a
multiplicative factor f$_{Sp}$ to take into account the possible
behavior of magnetic fields below the resolution limit of the
simulation.

>>>>

\enzo\ implements the equations of isotropic heat conduction in a manner
similar to that of \citet{2007ApJ...664..135P}.  
The isotropic flux of heat is given by:
\begin{equation}
\myvec{F} = -\kappa_{sp} \nabla T
\end{equation}
where $\kappa_{sp}$ is the Spitzer conduction coefficient, $4.6 \times 10^{-7}$~T$^{5/2}$ erg
s$^{-1}$~cm$^{-1}$~K$^{-1}$ \citep{1962pfig.book.....S}, here we are using a
value for the Coulomb logarithm, $\log \Lambda = 37.8$, that is
appropriate for the intracluster medium \citep{1988xrec.book.....S}.
It is quite possible that the local heat flux computed in this way 
can become unphysically large in
the high-temperature, low-density cluster regime when using this
formulation; therefore, we take into account the saturation of the heat
flux \citep{1977ApJ...211..135C} at a maximum level of
\begin{equation}
F_{sat} \simeq 0.4 n_e k_b T \left( \frac{2 k_b T}{\pi m_e} \right)^{1/2}.
\end{equation}
To ensure a smooth transition between
the Spitzer and saturated regimes, we define an effective
conductivity using the formalism of \citet{1988xrec.book.....S}
\begin{equation}
\kappa_{eff} = \frac{\kappa_{sp}}{1 + 4.2 \lambda_e / \ell_T} \; ,
\end{equation}
where $\lambda_e$ is the electron mean free path and $\ell_T \equiv T
/ |\nabla T|$ is the characteristic length scale of the local
temperature gradient.  We also assume that the conductivity of the
plasma can be described in terms of an effective conductivity, which
can be expressed as a fraction f$_{Sp}$ of the Spitzer conductivity
(where f$_{Sp} \leq 1.0$ are considered physical values).  This takes
into account physical processes below the resolution limit of the
simulation (such as tangled magnetic fields) that can suppress heat
transport.

Thermal conduction in a plasma can be strongly affected by the
presence of magnetic field lines, which may suppress heat
flow perpendicular to the magnetic field.  In that case, we allow for
heat transport only parallel to the magnetic field lines in
magnetohydrodynamic simulations.  Mathematically, this is treated as:
\begin{equation}
\myvec{F} = -\kappa_{sp} \hat{b} (\hat{b} \cdot \nabla T)
\end{equation}
where $\hat{b}$ is the unit vector denoting the direction of the
magnetic field.  As with the isotropic thermal conduction, we allow a
multiplicative factor f$_{Sp}$ to take into account the possible
behavior of magnetic fields below the resolution limit of the
simulation.
